---
title: "Analysis of Income and Market Value"
author: "Eric, Visruth, Sebastian, Ryan"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
    number-sections: true
    code-fold: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

# Analysis of Income and Market Value

In this project, we analyze the relationship between the daily income of individuals and the market value of listed companies (as a percentage of GDP) across various countries. The data for this analysis comes from two CSV files: `market_value_of_listed_companies_percent_of_gdp.csv` and `mincpcap_cppp.csv`.

## Data and Variable Description

### Data Sources

1.  **Market Value of Listed Companies (% of GDP)**

    -   This dataset provides information on the market value of all publicly traded companies in various countries, expressed as a percentage of the country's GDP for multiple years. The market value of listed companies refers to the total value of a company's shares of stock. It is calculated by multiplying the share price by the number of outstanding shares. This value represents the aggregate value of the companies' equity on the stock market.

    -   **Variables**

        -   `country`: The name of the country.

        -   `year`: The year for which the data was recorded.

        -   `market_value`: The market value of listed companies as a percentage of GDP.

2.  **Daily Income**

    -   This dataset provides the average daily income per capita for various countries over multiple years, adjusted for purchasing power parity (PPP). PPP is an economic theory that compares different countries' currencies through a basket of goods approach. It allows for more accurate comparisons of living standards and economic productivity between countries by accounting for differences in price levels.

    -   **Variables**

        -   `country`: The name of the country.

        -   `year`: The year for which the data is recorded.

        -   `income`: The average daily income per capita in constant PPP dollars.

### Data Collection Methods

The data on the market value of listed companies and daily income per capita is collected by various international financial organizations and statistical agencies. These organizations compile data from national statistical offices, stock exchanges, and other financial institutions. The market value data is derived from stock market data, while the income data is adjusted for PPP to account for differences in cost of living across countries.

### Hypothesized Relationship

We hypothesize that there is a positive relationship between a country's average daily income per capita and the market value of its listed companies as a percentage of GDP. This hypothesis is based on the idea that higher individual incomes could correlate with higher levels of investment, corporate profits, and economic activity, leading to greater market value of listed companies.

```{r}
library(tidyverse)
library(broom)
library(gganimate)
library(gifski)
library(knitr)
library(kableExtra)
library(png)

set.seed(0)

market <- read_csv("market_value_of_listed_companies_percent_of_gdp.csv")
income <- read_csv("mincpcap_cppp.csv")
```

## Data Cleaning Process

### Importing and Reshaping Data

First, we import the data from the two CSV files. The data is initially in a wide format, where each year is represented as a separate column. To facilitate analysis, the data is reshaped into a long format, where each row represents a single country's data for a specific year. This restructuring helps in merging the two datasets and for analysis.

### Merging Datasets

Once the data is reshaped, we merge the two datasets based on the common columns, which are the country names and years.

### Handling Missing Values

Any rows with missing values in either the market value or income columns are removed. This step is crucial as missing data can lead to biased results and affect the validity of the analysis.

### Summary of Cleaned Data

The resulting dataset contains the following columns:

-   `country`: The name of the country.

-   `year`: The year for which the data is recorded.

-   `market_value`: The market value of listed companies as a percentage of GDP.

-   `income`: The average daily income per capita in constant PPP dollars.

```{r}
market <- market |>
  pivot_longer(cols = -country, names_to = "year", values_to = "market_value")

income <- income |>
  pivot_longer(cols = -country, names_to = "year", values_to = "income")

data <- market |>
  inner_join(income, by = c("country", "year")) |>
  drop_na() # removing years which don't have both market value and income data
```

By cleaning and merging the datasets, we have prepared a comprehensive dataset that allows us to explore the relationship between daily income and the market value of listed companies across different countries and years. The next steps involve performing statistical analyses and visualizations to test our hypothesis and uncover insights from the data.

# Linear Regression

## Data Visualization

### Visualize the relationship between the quantitative variables

In order to understand the relationship between income and the market value (% of GDP), we will start by graphing a scatterplot of the two variables.

```{r}
data |>
  ggplot(aes(
    x = income,
    y = market_value
  )) +
  geom_jitter() +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(
    title = "Scatterplot of Income and Market Value",
    y = "",
    subtitle = "Market Value",
    x = "Income"
  )
```

The scatterplot shows that the majority of the data points are concentrated below the market value of 500, indicating that most observed market values are relatively low. There are a few data points with significantly higher market values, forming a sparse distribution above 500, which are identified as outliers. The scatterplot includes a linear regression line with a positive slope, suggesting a slight positive correlation between Income and Market Value. As income increases, market value tends to increase linearly, though the effect is small.

### Visualize the relationship between the quantitative variables

To further examine how these variables have evolved over time, we created an animated plot which shows how the relationship between income and the market value (% of GDP) of each country have changed each year.

```{r}
data |>
  mutate(year = as.integer(year)) |>
  ggplot(aes(
    x = income,
    y = market_value,
    color = country
  )) +
  geom_jitter() +
  theme_classic() +
  labs(
    title = "Scatterplot of Income and Market Value",
    y = "Market Value",
    x = "Income",
    caption = "(color represents each country, helps to track trajectories of individual countries)"
  ) +
  theme(legend.position = "none") +
  transition_time(year) +
  labs(subtitle = "Year: {frame_time}")
```

The animated plot significantly enhances the understanding of the linear relationship between the two variables. As the animation progresses, it becomes evident that as the income of countries increases over the years, there is a general trend of increasing market value (measured as a \$ of GDP) for public companies. Furthermore, the color of each data point is mapped to specific countries, making it easier to track and follow the trajectories of individual countries. This is particularly useful in identifying patterns and deviations from the overall trend. For instance, the outlier points identified in the initial static scatterplot are now shown to originate from the same country. This revelation provides context to the outliers, demonstrating that they are not indicative of the general relationship observed in most other countries.

## Linear Regression Model (LRM)

In this section, we continue our analysis of the relationship between income and the market value (% of GDP) by employing a linear regression model.

### What's an LRM?

Linear regression is a statistical technique used to model the relationship between two variables: a dependent variable (in this case, the mean market value of listed companies as a percentage of GDP) and one or more independent variables (here, the mean income per capita). It assumes a linear relationship, meaning that a change in the independent variable is associated with a proportional change in the dependent variable.

However, before we can do this, we first must start by condensing our data to be more accurately represented by the regression model. We do this by finding the average income and market value of each country. Below is a scatterplot of the summarized data.

```{r}
data_means <- data |>
  group_by(country) |> # grouping by country
  summarize(
    mean_market_value = mean(market_value),
    sem_market_value = sd(market_value) / sqrt(n()),
    mean_income = mean(income),
    sem_income = sd(income) / sqrt(n()),
    .groups = "drop"
  )
```

```{r}
data_means |>
  ggplot(aes(
    x = mean_income,
    y = mean_market_value
  )) +
  geom_jitter() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(
    title = "Scatterplot of Mean Country Income and Mean Country Market Value",
    subtitle = "Market Value",
    y = "",
    x = "Income"
  )
```

Now that the data has been summarized, we move on to fitting the linear model to our data. The table below shows the relationship between the two variables and will be used to construct the estimated regression model.

```{r}
mv_lm <- lm(mean_market_value ~ mean_income, data = data_means)
mv_lm_summary <- broom::tidy(mv_lm)



knitr::kable(mv_lm_summary,
  col.names = gsub(
    "[.]",
    " ",
    c("Term", "Coefficient Estimate", "Standard Error", "T-Statistic", "P-Value")
  ),
  align = "ccccc",
  caption = "Summary of Linear Regression Model Coefficients",
  digits = 3
)
```

In general, an estimated regression model can be represented as:

$$
\text{Market Value (\% of GDP)} = \hat{\beta}_0 + \hat{\beta}_1 \times \text{Daily Income per Capita} + \epsilon  
$$

Using the estimated coefficent values from the regression model, we can fill in the value of $\hat{\beta}_0$, the y-intercept, and the value of $\hat{\beta}_1$, the scaler which defines the relationship between the explainatory (income) and response variable (market value as a percent of GDP). The resulting equation is defined below:

$$
\text{Market Value (\% of GDP)} = 22.25 + 1.025 \times \text{Daily Income per Capita} + \epsilon
$$ The interpretation of this equation is that for any increase in daily income per capita, market value (% of GDP) will increase by 0.025 plus 22.25 percent.

### Interpretation of the coefficients:

This intercept (22.25) suggests that even with zero mean income per capita, the market value of listed companies would be estimated to be 22.25% of GDP. This interpretation should be taken with caution as it represents an extrapolation beyond the observed data.

The slope (1.025) indicates that for every \$1,000 increase in mean income per capita, the market value of listed companies is estimated to increase by 1.025% of GDP, on average. This suggests a positive association between income and market value.

## Model Fit

```{r}
mv_lm_augmented <- augment(mv_lm)

# Calculate Variances
response_var <- var(mv_lm_augmented$.fitted +
  mv_lm_augmented$.resid)

fitted_var <- var(mv_lm_augmented$.fitted)

residual_var <- var(mv_lm_augmented$.resid)

# Create a DF for the Table
variance_table <- data.frame(
  Source = c(
    "Response Values",
    "Fitted Values",
    "Residuals"
  ),
  Variance = c(
    response_var,
    fitted_var,
    residual_var
  )
)

# Format with kableExtra
kableExtra::kable(
  variance_table,
  align = "cc",
  col.names = c(
    "Source of Variation",
    "Variance"
  ),
  digits = 3,
  caption = "Table of Variances from the Linear Regression Model",
  booktabs = TRUE
) |>
  kable_styling(latex_options = "striped", full_width = FALSE)
```

The variance table reveals that the fitted values from our model account for only 13% (450.531 / 3466.107) of the total variation in market values. This indicates that the majority of the fluctuation in market value is due to factors other than mean income per capita. While a statistically significant positive relationship exists, this model alone has limited ability to predict market values based solely on income.

```{r, r-squared}
r_sq <- fitted_var / response_var

value_r_sq <- paste("R-squared:", round(r_sq, 3))

value_r_sq
```

The R-squared value suggests that only approximately 13% of the variation in market value is explained by income. This implies that other factors not included in our model play a significant role in determining market value.

# Simulation
## Visualizing Simulations from the Model
#### Generating Predictions and Simulated Data

First, we used a simple linear regression model to generate predictions for the market value based on income. These predictions were then adjusted with random errors based on the residual standard error from the linear regression model to create simulated data. The resulting scatterplots, both for the observed and simulated data, were analyzed to understand the similarities and differences between them.

```{r simulated data}
#| output: false
pred <- predict(mv_lm)
est_sigma <- sigma(mv_lm)
```

```{r}
rand_error <- function(x, mean = 0, sd){
  x + rnorm(length(x), mean = mean, sd = sd)
}

sim_response <- tibble(sim_mv = rand_error(pred,
                                            sd = est_sigma))
```

```{r}
data_means |>
  bind_cols(sim_response) |>
  ggplot(aes(
    x = mean_income,
    y = sim_mv
  )) +
  geom_jitter() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(
    title = "Scatterplot of Mean Country Income and Mean Country Market Value",
    subtitle = "Simulated Market Value",
    y = "",
    x = "Income"
  )
```

#### Comparing Simulated Data to Observed Data

Both scatterplots show a positive trend line, confirming our hypothesis that higher income correlates with higher market value. In both plots, a significant number of data points are below a market value of 100, indicating that most countries have market values within this range. Additionally, both scatterplots contain outliers. This suggests variability in market values that the model attempts to capture.

The observed data has a more skewed distribution with extreme outliers exceeding 500, while the simulated data points are more evenly spread around the regression line, displaying a more normal distribution. The observed data shows higher variability and greater spread in market values, indicating that factors other than income significantly influence market value. On the other hand, the simulated data shows a more constrained spread, emphasizing the model’s limitations in capturing the variability in market values.

The comparison underscores that while the linear regression model depicts the general positive trend between income and market value, it underestimates the variability observed in the real-world data. The outliers in the observed data suggests that additional factors beyond income play a role in determining the market value of listed companies.

## Generating Multiple Predictive Checks
The purpose of generating simulated datasets is to provide a benchmark against which we can compare our observed data. By comparing the observed data to the distribution of simulated data, we can assess whether our model's assumptions are reasonable and how well it captures the underlying data generating process.
```{r, Generating Sim Data}
n_sims <- 1000
sim_data <- matrix(NA, nrow = nrow(data_means), ncol = n_sims)

for (i in 1:n_sims) {
  sim_data[, i] <- rand_error(pred, sd = est_sigma)
}

sim_data <- as.data.frame(sim_data)
names(sim_data) <- paste0("sim", 1:n_sims)
sim_data <- cbind(data_means, sim_data)
```


### Comparing Against Regression
To assess how well our linear model captures the true relationship between mean income and market value, we conducted predictive checks. These checks involve generating numerous simulated datasets based on our model's parameters and comparing them to our observed data. Each simulated dataset represents a plausible outcome if our model were a perfect representation of reality.
```{r}
# vector to store values
r_squared_values <- numeric(n_sims)
for (i in 1:n_sims) {
  
  sim_data_temp <- sim_data |> 
    select(mean_income, starts_with("sim")) |> 
    rename(simulated_mv = paste0("sim", i)) # Rename 
  
  # Join
  sim_data_temp <- sim_data_temp |> 
    inner_join(data_means|> select(mean_income, mean_market_value), by = "mean_income")

  # LRM
  sim_lm <- lm(mean_market_value ~ simulated_mv, data = sim_data_temp)
  
  # Store
  r_squared_values[i] <- summary(sim_lm)$r.squared
}
```
We regressed the observed market values against each of the 1,000 simulated datasets, calculating the R-squared value for each regression. The distribution of these R-squared values reveals the range of model fit we would expect to see if our model were accurate.
### Plotting R-squared Distribution
Visualize distribution of the R-sq. values.
```{r}
# Create a histogram of the R-squared values
ggplot(data.frame(R_squared = r_squared_values), 
       aes(x = R_squared)) +
  geom_histogram(binwidth = 0.02,
                 fill = "skyblue",
                 color = "black",
                 alpha = 0.7) +
  
  labs(title = "Distribution of R-squared Values from Predictive Checks",
       x = "R-squared",
       y = "Frequency") +
  theme_minimal()
```

**Implications:**

+  **Distribution of R-squared:**
The histogram shows the distribution of R-squared values we would expect to see if our model were a good fit. The observed R-squared from our model on the real data (0.13) falls within the lower range of the simulated R-squared values.
    
    
+  **Model Adequacy:**
This suggests that while our model isn't a terrible fit, it's also not exceptionally good at generating data similar to what was observed. There is a considerable chance that we could observe data that has a better fit (higher R-squared) with the model than our original data. This further reinforces the idea that there are likely other factors influencing market value that our model isn't capturing.

**Considerations:**

+   **Assumption of Correct Model:**
The simulations are based on the assumption that the linear model is the correct model for the data. If the underlying relationship is different (e.g., non-linear), the predictive checks might not be reliable.


+  **Data Limitations:**
The quality of the simulations depends on the quality of the original data. If there are measurement errors or outliers in the original data, these will also affect the simulations.