---
title: "Unraveling the Link Between Income and Market Value: A Global Perspective"
format: html
editor: visual
embed-resources: true
code-tools: true
code-fold: true
---

# Introduction

The relationship between a nation's wealth and the performance of its stock market is a subject of ongoing economic inquiry. In this study, we delve into this relationship, specifically investigating the link between average daily income per capita and the market value of listed companies as a percentage of Gross Domestic Product (GDP). Our analysis draws upon a comprehensive dataset spanning various countries and multiple years, allowing us to explore this relationship across diverse economic contexts.

We hypothesize that a positive correlation exists between daily income per capita and the market value of listed companies. This hypothesis stems from the notion that higher individual incomes may fuel increased investment, bolster corporate profits, and stimulate overall economic activity, ultimately contributing to a greater market value for publicly traded companies.

To test this hypothesis, we employ a linear regression model and conduct rigorous predictive checks to assess the model's adequacy. Our approach encompasses data cleaning, exploratory data analysis, model fitting, and validation, ensuring a comprehensive investigation of the proposed relationship.

## Data Sources and Preparation

Our analysis draws upon data from two primary sources:

1.  **Market Value of Listed Companies (% of GDP):** This dataset, compiled by international financial organizations, provides the market value of all publicly traded companies in a given country, expressed as a percentage of its GDP for various years.

2.  **Daily Income:** This dataset, gathered from statistical agencies, reports the average daily income per capita across different countries over multiple years, adjusted for purchasing power parity (PPP) to account for variations in living costs.

We begin by reshaping the data from wide to long format, merging the two datasets based on country and year, and removing any observations with missing values. This ensures a clean and complete dataset for our subsequent analysis.

```{r, setup}
#| label: setup
#| message: false
#| warning: false
#| echo: false

library(tidyverse)
library(broom)
library(gganimate)
library(gifski)
library(knitr)
library(kableExtra)
library(png)

set.seed(0)

market <- read_csv("market_value_of_listed_companies_percent_of_gdp.csv")
income <- read_csv("mincpcap_cppp.csv")
```

```{r}
#| label: clean-and-merge
#| message: false
#| warning: false
#| echo: false

# Reshape to long format
market_long <- market |> 
  pivot_longer(!country, names_to = "year", values_to = "market_value")

income_long <- income |> 
  pivot_longer(!country, names_to = "year", values_to = "income")

# Convert year column to numeric
market_long$year <- as.numeric(market_long$year)
income_long$year <- as.numeric(income_long$year)

# Merge datasets
data <- inner_join(market_long, income_long, by = c("country", "year"))

# Remove rows with missing values
data <- data |> 
  drop_na()
```
